\documentclass[11pt]{article}
\usepackage{xunicode}
\usepackage{fontspec}
\usepackage{xltxtra}
\usepackage{float}
\usepackage{fancyvrb}
\defaultfontfeatures{Scale=MatchLowercase}
\setmainfont[Mapping=tex-text,Ligatures=TeX,Numbers=OldStyle]{Linux Libertine O}
\setsansfont[Mapping=tex−text,Numbers=OldStyle] {Linux Biolinum O}
\setmonofont[Mapping=tex-text,Scale=0.8] {DejaVu Sans Mono}

\newcommand{\timeflies}{\emph{TimeFlies}}

%\setromanfont{Linux Libertine O}

\begin{document}

\DefineVerbatimEnvironment{inputfile}{Verbatim}
   {%frame=single,
    %framesep=1mm,
    numbers=left,
    baselinestretch=0.8,
    labelposition=topline}

%\newfloat{example}{thp}{lof}[section]
%\floatname{example}{Example}

\title{\timeflies\ -- A Tool for Time Keeping}
\author{J\"org Bullmann}
\maketitle
\tableofcontents
\newpage

\section{What is \timeflies\ About?}

Do you need to account for the time you spend at work? What project or work package have you been working on? Do you want to keep track of the hours you work? How much leave have you got left for this year?

Do you keep a daily work log containing things you did, problems you solved, some kind of to do list?

Do you want to make an estimate of effort for a project or work package? Would you like to break down those things into smaller items and possibly break down those again too?

\timeflies\ can help you with this.

\section{Tutorial by Example}

In this section we will look at a number of use cases. All \timeflies\ data is kept in plain text files. So all your data is always easily accessible to you and the format itself is quite human-readable. Moreover it can easily be version controlled.

\subsection{Recording Time}

To record your work time keep a work log file with \verb-day- lines specifying the dates and in and out times telling when you arrived at work and when you left. The times can be given in decimals or in hours and minutes:

\input{fly-1.in}

Do you need to account for breaks you are taking? Use \verb-off- instructions to state periods of time in a day of work during which you were not actually working:

\input{fly-2.in}

This last file is equivalent to the following:

\input{fly-3.in}

This last example illustrates the notion of a \verb-day-\emph{-block}: a \verb-day-\emph{-block} extends
from one \verb-day--keyword to the next and everything inside this day block is part of that day.

The days in the file do not need to be listed chronologically. You could e.g.\ list the days in reverse order so that the present is always at the top of the file.

If you want to mask out part of your log temporarily you can use the \verb-#- source comment marker. \timeflies\ ignores the \verb-#- and everything following it until the end of line. It works just the same as e.g.\ a Python comment.

\input{fly-4.in}

\subsection{Keeping Notes in the Log}

Do you want to keep notes about your work in the same place as you keep the time information? Use \emph{log comment} lines like in this file:

\input{fly-5.in}

A log comment line starts in double dash characters \verb=--= and one or more space characters. All text following these characters until the end of the line (or until a \verb-#- source comment marker) with trailing spaces removed constitute the recorded log comment.

Now what can you do with such a file? Assume the above work log file's name is \verb:work-log.fly:, then option \verb:-t: tells \timeflies\ to calculate your work times.

\input{fly-5a.out}

To include the log comments in this output, use the \verb=-C= option:

\input{fly-5b.out}

\subsection{Time Summaries}

Here's a longer example where you can see the use of weekly and montly summaries.

\input{fly-8.in}

Use option \verb:-f week: to get an overview of weekly work time balances.

\input{fly-8a.out}

Or have both weekly and monthly balances shown.

\input{fly-8b.out}

Maybe you only want to look at one month?

\input{fly-8c.out}

You only need weekly totals in that one month?

\input{fly-8d.out}

\subsection{Logging Activities}

Log comments are a good way to keep track of things you don't want to forget and have accessible and also aligned with your work time line. Log comments have no work effort assigned to them, though. So you cannot use them in any way for calculations of effort spent.

You use \emph{work packages} and \emph{activities} to connect the time you work with the work packages you work on: first, you define your work packages, then you use \emph{activity} lines in the day blocks instead of log comment lines.

A work package definition is a line starting with the keyword \verb:work-package: (or its abbreviation \verb:wp:) followed by a work package name.

An activity line starts with a single dash character \verb:-: followed by one or more spaces. This is followed by a work package id and a duration. This is optionally followed by a comma and some activity comment.

See below the converted example work log file.

\input{fly-6.in}

Option \verb:-w: tells \timeflies\ to calculate the times you have been working on the different work packages:

\input{fly-6a.out}

To also show the activities contributing to the different work packages, use option \verb:-a::

\input{fly-6b.out}

To check whether you have allocated all your working time to work packages, use option \verb:-c::

\input{fly-6c.out}

This shows that on two days the time at work and the time worked on work packages are differing.

\subsection{Work Package Breakdown}

In the previous section, work packages have been defined as simple, atomic, named items. A work package can be subdivided and refined hierarchically. See the following example.

\input{fly-7-wp.in}

The items in this work package hierarchy can be referred to in activity lines as \verb:.:-delimited work package path names.

Following, a piece of work log for the above project.

\input{fly-7.in}

Assuming the work package defintions are kept in file \verb:prj-mighty-digester.fly: and the work log itself in \verb:work-log.fly:, the work package summary can be calculated as follows with option \verb:-w: as in the previous example.

\input{fly-7a.out}

And here the same with activities shown.

\input{fly-7b.out}

%\subsection{Monitor Project Progress}

\section{Reference}

\subsection{Command Line Options}

\subsection{File Syntax}

\end{document}
